IMPC Mouse data - Variance in sex differences

### 2019-08-03 scripts/2019_July_IMPC variance and sex difference.Rmd
clean up

---
title: "IMPC Mouse data - Variance in sex differences"
author: "Susanne & Felix Zajitschek"
date: "July 2019"
output:
  html_document:
    code_folding: hide
    toc: true
    toc_float: true
    toc_depth: 3
runtime: shiny
---

@@

We here provide the cleaned dataset, which we have saved in a folder called "export", as easy starting point. However, the full dataset can be loaded and cleaned using the data cleaning function (Function 1 above), if "#" signs in the code below are removed  (created as that is much smaller than the .csv - which we can still provide for those who absolutely want to start from scratch?)
```{r}
## Load raw data - save cleaned dataset as RDS for reuse
#data_raw <- load_raw("data/dr7.0_all_control_data.csv") %>%
#    clean_raw_data()
#dir.create("export", F, F)
#saveRDS(data_raw, "export/data_clean.rds")

data1 <- readRDS("export/data_clean.rds")
#data1 
```

@@

 group <- read.csv("export/ParameterGrouping.csv") 
 data <- data3
 data$parameterGroup <- group$parameter[match(data$parameter_name, group$parameter_name)]

@@

trait_meta_results <- write.csv(results.alltraits.grouping, file = "export/trait_meta_results.csv")

@@

```{r}
procedures <- read.csv("export/procedures.csv")

@@

trait_meta_results <- write.csv(metacombo, file = "export/trait_meta_results.csv")
```

@@

# * Heterogeneity
# ** FIGURE 4 (Second-order meta analysis on heterogeneity)
## *Create matrix to store results for all traits
```{r}
results.allhetero.grouping <- as.data.frame(cbind(c(1:n), matrix(rep(0, n*30), ncol = 30))) 
names(results.allhetero.grouping) <- c("id", "sigma2_strain.CVR", "sigma2_center.CVR", "sigma2_error.CVR", "s.nlevels.strain.CVR", 
	"s.nlevels.center.CVR", "s.nlevels.error.CVR", "sigma2_strain.VR", "sigma2_center.VR", "sigma2_error.VR", "s.nlevels.strain.VR", 
	"s.nlevels.center.VR", "s.nlevels.error.VR", "sigma2_strain.RR", "sigma2_center.RR", "sigma2_error.RR", "s.nlevels.strain.RR", 
	"s.nlevels.center.RR", "s.nlevels.error.RR", "lnCVR", "lnCVR_lower", "lnCVR_upper", "lnCVR_se", "lnVR", "lnVR_lower", "lnVR_upper", 
	"lnVR_se", "lnRR", "lnRR_lower", "lnRR_upper" ,"lnRR_se")

@@

# **LOOP
#  parameters to extract from metafor (sigma2's, s.nlevels)

@@

## *Exclude traits
```{r}
results.allhetero.grouping2 <- results.allhetero.grouping[results.allhetero.grouping$s.nlevels.strain.VR != 0, ]
nrow(results.allhetero.grouping2) #218
```

# *Merge data sets containing metafor results with procedure etc. names 
```{r}
#procedures <- read.csv("procedures.csv")

@@

## *Correlated parameters

```{r}
metahetero1 <- results.allhetero.grouping2 
length(unique(metahetero1$procedure)) #18
length(unique(metahetero1$GroupingTerm)) #9 
length(unique(metahetero1$parameter_group)) # 149 levels: one more tha in effect size alanlysis, see above; CHECK!
length(unique(metahetero1$parameter_name)) #218

# **Count of number of parameter names (correlated sub-traits) in each parameter group (par_group_size)

```{r}
metahetero1b <-
  metahetero1 %>%
  group_by(parameter_group) %>% 
  mutate(par_group_size = n_distinct(parameter_name))

@@

# **Create subsets with > 1 count (par_group_size > 1)

```{r}
metahetero1_sub <- subset(metahetero1, par_group_size > 1) # 90 observations  
str(metahetero1_sub)
# metahetero1_sub$sampleSize <- as.numeric(metahetero1_sub$sampleSize) #from previous analysis? don't think is used: : delete in final version
```

# **Nest data
```{r}
n_count. <- metahetero1_sub %>% 
  group_by(parameter_group) %>% 
  #mutate(raw_N = sum(sampleSize)) %>%  #don't think is necessary: delete in final version
  nest()

```{r}
model_count. <- n_count. %>% 
  mutate(model_lnRR = map(data, ~ robu(.x$lnRR ~ 1, data= .x, studynum= .x$id, modelweights = c("CORR"), rho = 0.8, 
                                                small = TRUE, var.eff.size= (.x$lnRR_se)^2 )),
  model_lnVR = map(data, ~ robu(.x$lnVR ~ 1, data= .x, studynum= .x$id, modelweights = c("CORR"), rho = 0.8, 
                                                small = TRUE, var.eff.size= (.x$lnVR_se)^2 )),
  model_lnCVR = map(data, ~ robu(.x$lnCVR ~ 1, data= .x, studynum= .x$id, modelweights = c("CORR"), rho = 0.8, 
                                                small = TRUE, var.eff.size= (.x$lnCVR_se)^2 )))

```
#Robumeta object details:
#str(model_count.$model_lnCVR[[1]])

@@

```{r}
robusub_RR. <- model_count. %>% 
  transmute(parameter_group, estimatelnRR = map(model_lnRR, count_fun.)) %>% 
  mutate(r = map(estimatelnRR, ~ data.frame(t(.)))) %>%
  unnest(r) %>%
  select(-estimatelnRR) %>%
  purrr::set_names(c("parameter_group","var.RR","N.RR"))

@@

## **Merge the two data sets (the new [robu_all] and the initial [uncorrelated sub-traits with count = 1]) 
##  In this step, we 	
1) put the N from robumeta and the  N from metafor (s.nlevels.error) together into the same columns (N.RR, N.VR, N.CVR)

@@

@@

# **Combine data 
# Step1 
```{r}
combinedmetahetero <- bind_rows(robu_all., metahetero_all)
#glimpse(combinedmetahetero)
```
# Steps 2&3
```{r}
metacombohetero <- combinedmetahetero
metacombohetero$counts <- metahetero1$par_group_size[match( metacombohetero$parameter_group, metahetero1$parameter_group)]
metacombohetero$procedure2 <-metahetero1$procedure[match( metacombohetero$parameter_group, metahetero1$parameter_group)]
metacombohetero$GroupingTerm2 <-metahetero1$GroupingTerm[match( metacombohetero$parameter_group, metahetero1$parameter_group)]

```
# **Clean-up and rename 
```{r}
metacombohetero <- metacombohetero[, c(1:7, 43:45)] 
names(metacombohetero)[9] <- "procedure" 
names(metacombohetero)[10] <- "GroupingTerm"

```{r}
#write.csv(metacombohetero, "metacombohetero.csv")

## *Last step: meta-meta-analysis of heterogeneity
```{r}
glimpse(metacombohetero)

@@

####################################################################################################################################
# *PLOTS

# **Restructure data for plotting 
```{r}
heterog3 <- gather(heterog2, parameter, value, c(heteroCVR, heteroVR, heteroRR), factor_key= TRUE)

@@

# *Plot FIGURE 4 (Second-order meta analysis on heterogeneity)

# **Plot Fig4 all traits
```{r}

>>>>

---
title: "IMPC Mouse data - Variance in sex differences"
author: "Susanne & Felix Zajitschek"
date: "July 2019"
output:
  html_document:
    code_folding: hide
    toc: true
    toc_float: true
    toc_depth: 3
  html_notebook: 
    toc: yes
#runtime: shiny
---

@@

We here provide the cleaned dataset, which we have saved in a folder called "export", as easy starting point. However, the full dataset can be loaded and cleaned using the data cleaning function (Function 1 above), if "#" signs in the code below are removed  (created as that is much smaller than the .csv - which we can still provide for those who absolutely want to start from scratch?)
```{r}
## Load raw data - save cleaned dataset as RDS for reuse
#data_raw <- load_raw("data/dr7.0_all_control_data.csv") %>%
#    clean_raw_data()
#dir.create("export", F, F)
#saveRDS(data_raw, "export/data_clean.rds")
getwd()
data1 <- readRDS("/export/data_clean.rds") 
#data1 
```

@@

 group <- read.csv("/export/ParameterGrouping.csv") 
 data <- data3
 data$parameterGroup <- group$parameter[match(data$parameter_name, group$parameter_name)]

@@

#trait_meta_results <- write.csv(results.alltraits.grouping, file = "export/trait_meta_results.csv")

@@

```{r}
procedures <- read.csv("/export/procedures.csv")

@@

trait_meta_results <- write.csv(metacombo, file = "/export/trait_meta_results.csv")
```

@@

# Heterogeneity
## FIGURE 4 (Second-order meta analysis on heterogeneity)
### Create matrix to store results for all traits
```{r}
results.allhetero.grouping <- as.data.frame(cbind(c(1:n), matrix(rep(0, n*30), ncol = 30))) 
names(results.allhetero.grouping) <- c("id", "sigma2_strain.CVR", "sigma2_center.CVR", "sigma2_error.CVR", "s.nlevels.strain.CVR", 
	"s.nlevels.center.CVR", "s.nlevels.error.CVR", "sigma2_strain.VR", "sigma2_center.VR", "sigma2_error.VR", "s.nlevels.strain.VR", 
	"s.nlevels.center.VR", "s.nlevels.error.VR", "sigma2_strain.RR", "sigma2_center.RR", "sigma2_error.RR", "s.nlevels.strain.RR", 
	"s.nlevels.center.RR", "s.nlevels.error.RR", "lnCVR", "lnCVR_lower", "lnCVR_upper", "lnCVR_se", "lnVR", "lnVR_lower", "lnVR_upper", 
	"lnVR_se", "lnRR", "lnRR_lower", "lnRR_upper" ,"lnRR_se")

@@

## LOOP
parameters to extract from metafor (sigma2's, s.nlevels)

@@

### Exclude traits
```{r}
results.allhetero.grouping2 <- results.allhetero.grouping[results.allhetero.grouping$s.nlevels.strain.VR != 0, ]
nrow(results.allhetero.grouping2) #218
```

Merge data sets containing metafor results with procedure etc. names 
```{r}
#procedures <- read.csv("procedures.csv")

@@

### Dealing with correlated parameters
```{r}
metahetero1 <- results.allhetero.grouping2 
length(unique(metahetero1$procedure)) #18
length(unique(metahetero1$GroupingTerm)) #9 
length(unique(metahetero1$parameter_group)) # 149 levels: one more tha in effect size alanlysis, see above; CHECK!
length(unique(metahetero1$parameter_name)) #218

# Count of number of parameter names (correlated sub-traits) in each parameter group (par_group_size)

metahetero1b <-
  metahetero1 %>%
  group_by(parameter_group) %>% 
  mutate(par_group_size = n_distinct(parameter_name))

@@

#Create subsets with > 1 count (par_group_size > 1)

metahetero1_sub <- subset(metahetero1, par_group_size > 1) # 90 observations  
str(metahetero1_sub)
# metahetero1_sub$sampleSize <- as.numeric(metahetero1_sub$sampleSize) #from previous analysis? don't think is used: : delete in final version

# Nest data

n_count. <- metahetero1_sub %>% 
  group_by(parameter_group) %>% 
  #mutate(raw_N = sum(sampleSize)) %>%  #don't think is necessary: delete in final version
  nest()

# meta-analysis preparation

model_count. <- n_count. %>% 
  mutate(model_lnRR = map(data, ~ robu(.x$lnRR ~ 1, data= .x, studynum= .x$id, modelweights = c("CORR"), rho = 0.8, 
                                                small = TRUE, var.eff.size= (.x$lnRR_se)^2 )),
  model_lnVR = map(data, ~ robu(.x$lnVR ~ 1, data= .x, studynum= .x$id, modelweights = c("CORR"), rho = 0.8, 
                                                small = TRUE, var.eff.size= (.x$lnVR_se)^2 )),
  model_lnCVR = map(data, ~ robu(.x$lnCVR ~ 1, data= .x, studynum= .x$id, modelweights = c("CORR"), rho = 0.8, 
                                                small = TRUE, var.eff.size= (.x$lnCVR_se)^2 )))

#Robumeta object details:
#str(model_count.$model_lnCVR[[1]])

@@

robusub_RR. <- model_count. %>% 
  transmute(parameter_group, estimatelnRR = map(model_lnRR, count_fun.)) %>% 
  mutate(r = map(estimatelnRR, ~ data.frame(t(.)))) %>%
  unnest(r) %>%
  select(-estimatelnRR) %>%
  purrr::set_names(c("parameter_group","var.RR","N.RR"))

@@

### **Merge the two data sets (the new [robu_all] and the initial [uncorrelated sub-traits with count = 1]) 
  In this step, we 	
1) put the N from robumeta and the  N from metafor (s.nlevels.error) together into the same columns (N.RR, N.VR, N.CVR)

@@

# **Combine data 
# Step1

combinedmetahetero <- bind_rows(robu_all., metahetero_all)
#glimpse(combinedmetahetero)

# Steps 2&3

metacombohetero <- combinedmetahetero
metacombohetero$counts <- metahetero1$par_group_size[match( metacombohetero$parameter_group, metahetero1$parameter_group)]
metacombohetero$procedure2 <-metahetero1$procedure[match( metacombohetero$parameter_group, metahetero1$parameter_group)]
metacombohetero$GroupingTerm2 <-metahetero1$GroupingTerm[match( metacombohetero$parameter_group, metahetero1$parameter_group)]

# **Clean-up and rename

metacombohetero <- metacombohetero[, c(1:7, 43:45)] 
names(metacombohetero)[9] <- "procedure" 
names(metacombohetero)[10] <- "GroupingTerm"

#write.csv(metacombohetero, "metacombohetero.csv")

@@

## *Last step: meta-meta-analysis of heterogeneity
```{r}
#glimpse(metacombohetero)

@@

####################################################################################################################################
## Heterogeneity PLOTS

Restructure data for plotting 
```{r}
heterog3 <- gather(heterog2, parameter, value, c(heteroCVR, heteroVR, heteroRR), factor_key= TRUE)

@@

## *Plot FIGURE 4 (Second-order meta analysis on heterogeneity)

Plot Fig4 all traits
```{r}
